preciseTAD: A machine learning framework for precise TAD boundary prediction at base pair resolution

### 2020-07-23 Manuscript/manuscript.Rmd
Add micro-C discussion

A limitation of *preciseTAD* is that it is dependent on "ground truth" boundaries provided by a user-defined TAD caller. Given the wide variery of TAD callers and their variable performance [@Zufferey:2018aa; @forcato2017comparison], it is hard to define "ground truth" TAD boundaries. Ideally we would benchmark *preciseTAD* against simulated boundaries. While methods for simulating Hi-C data sets wish boundary annotations exist [@forcato2017comparison; @zheng2020freehi], methods for simulating the associated genomic annotations (the main component guiding *preciseTAD* predictions) are lacking. Given *preciseTAD* can be trained on the whole-genome data, we feel that the total number of TAD/loop boundaries is sufficient to guide learning of the association between genomic annotations and boundaries for precise boundary predictions.

Another limitation of *preciseTAD* is that it does not provide hierarchical boundaries, but instead provides the coordinates of boundary points and regions. Studies have shown evidence of the hierarchical nature to the 3D architecture of the human genome [@hansen2018recent; @weinreb2016identification; @fraser2015hierarchical]. In such a hierarchy, the boundaries provided by *preciseTAD* can be interpreted as the lowest level of boundaries [??? This can be debated. We are training on and predicting certain level of boundaries. Suggesting discussing that false positives detected by preciseTAD may represent boundaries at finer resolution].

>>>>

A limitation of *preciseTAD* is that it is dependent on "ground truth" boundaries provided by a user-defined TAD caller. Given the wide variery of TAD callers and their variable performance [@Zufferey:2018aa; @forcato2017comparison], it is hard to define "ground truth" TAD boundaries. <!--Ideally we would benchmark *preciseTAD* against simulated boundaries. While methods for simulating Hi-C data sets with boundary annotations exist [@forcato2017comparison; @zheng2020freehi], methods for simulating the associated genomic annotations (the main component guiding *preciseTAD* predictions) are lacking.-->Ultra-deep Hi-C sequencing [@bonev2017multiscale] and newer technologies for precise mapping of chromatin interactions (e.g., Micro-C [@Krietenstein:2020aa]), coupled with more precise technologies for genomic annotation profiling (e.g., CUT&RUN for precise mapping of transcription factor binding sites) will help to refine the location and the genomic signatures of "ground truth" TAD boundaries. In the current work, given *preciseTAD* can be trained on the whole-genome data, we feel that the total number of TAD/loop boundaries is sufficient to guide learning of the association between genomic annotations and boundaries for precise boundary predictions.

Another limitation of *preciseTAD* is that it does not provide information about hierarchy of TAD boundaries, but instead provides the coordinates of boundary points and regions. Studies have shown evidence of the hierarchical nature to the 3D architecture of the human genome [@hansen2018recent; @weinreb2016identification; @fraser2015hierarchical]. In such a hierarchy, the boundaries provided by *preciseTAD* can be interpreted as the lowest level of boundaries [??? This can be debated. We are training on and predicting certain level of boundaries. Suggesting discussing that false positives detected by preciseTAD may represent boundaries at finer resolution].
